### 2016-11-20 paper.tex
Abstract

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for Web of Conferences Journal
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%%%% EDP Science %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%\documentclass[option comma separated list]{webofc}
%%%Three important options:
%%% "epj" for EPJ Web of Conferences Journal
%%% "bio" for BIO Web of Conferences Journal
%%% "mat" for MATEC Web of Conferences Journal
%%% "itm" for ITM Web of Conferences Journal
%%% "e3s" for E3S Web of Conferences Journal
%%% "shs" for SHS Web of Conferences Journal
%%% "twocolumn" for typesetting an article in two columns format (default one column)
\documentclass[epj,twocolumn]{webofc}
\usepackage[varg]{txfonts}   % Web of Conferences font
%
% Put here some packages required or/and some personnal commands
%
% Important: please activate and fill the "wocname" command with the exact title of the series for conferences not included in any of the series listed on the top
%
%\wocname{?????????}
%
% Very important: please fill the "woctitle" command with the exact title of the conference
%
\woctitle{Powders \& Grains 2017}
%
%
\begin{document}
%
\title{Collapse of tall granular columns in fluid}
%
% subtitle is optionnal
%
%%%\subtitle{Do you have a subtitle?\\ If so, write it here}

\author{\firstname{Krishna} \lastname{Kumar}\inst{1}\fnsep\thanks{\email{kks32@cam.ac.uk}} \and
        \firstname{Kenichi} \lastname{Soga}\inst{2}\fnsep\thanks{\email{soga@berkeley.edu}} \and
        \firstname{Jean-Yves} \lastname{Delenne}\inst{3}\fnsep\thanks{\email{}} \and
        \firstname{Farhang} \lastname{Radjai}\inst{4}\fnsep\thanks{\email{}}
        % etc.
}

\institute{Department of Engineering, University of Cambridge, UK 
\and
           University of California, Berkeley, US
\and
           INRA, France
\and 
           CNRS, France
          }

\abstract{%
  Avalanches, landslides, and debris flows are geophysical hazards, which involve rapid mass movement
  of granular solids, water, and air as a single phase system. In order to describe the mechanism of
  immersed granular flows, it is important to consider both the dynamics of the solid phase and the
  role of the ambient fluid (Denlinger and Iverson, 2001). The dynamics of the solid phase alone are
  insufficient to describe the mechanism of granular flows in fluid. It is important to consider the
  effect of hydrodynamic forces that reduce the weight of the solids inducing a transition from
  dense-compacted to dense-suspended flows, and the drag interactions which counteract the movement
  of the solids (Meruane et al., 2010). Transient regimes characterised by a change in the solid
  fraction, dilation at the onset of flow and the development of excess pore-pressure, result in
  altering the balance between the stress carried by the fluid and that carried by the grains, thereby
  changing the overall behaviour of the flow. In the present study, a two-dimensional coupled Lattice
  Boltzmann LBM – DEM technique is developed to understand the micro-scale rheology of granular
  flows in fluid.

  In the present study, the collapse of a granular column in fluid is studied using 2D LBM - DEM.
  The effect of initial aspect ratio on the run-out behaviour is investigated. The flow kinematics are
  compared with the dry and buoyant granular collapse to understand the influence of hydrodynamic forces
  and lubrication on the run-out. In the case of tall columns, the amount of material destabilised above
  the failure plane is larger than that of short columns. Hence in tall columns, the surface area of the
  mobilised mass that interacts with the surrounding fluid is significantly higher than the short columns.
  This increase in the area of soil - fluid interaction results in an increase in the formation of
  turbulent vortices that alter the deposit morphology during the collapse. It is observed that the
  vortices result in formation of heaps that significantly affect the distribution of mass in the flow.
  Staron and Hinch (2007) observed that the distribution of mass in a granular flow plays a crucial role
  in the flow kinematics. In order to understand the behaviour of tall columns, the run-out behaviour of a
  dense granular column with an initial aspect ratio of 6 is studied. The collapse of a tall granular
  column on slopes of 0, 2.5, 5 and 7.5 are studied. The permeability of the granular mass was varied
  to understand the effect of permeability on the run-out behaviour. }
%
\maketitle
%
\section{Introduction}
The collapse of a granular column on a horizontal surface is a simple case of granular flow, however a
proper model that describes the flow dynamics is still lacking. Experimental investigations have shown that
the flow duration, the spreading velocity, the final extent of the deposit, and the energy dissipation can be scaled in a quantitative way independent of substrate properties, grain size, density, and shape of the granular
material and released mass (Lajeunesse, Monnier, \& Homsy, 2005; Lube, Huppert, Sparks, \& Freundt, 2005).
Simple mathematical models based on conservation of horizontal momentum capture the scaling laws of the
final deposit, but fail to describe the initial transition regime. Granular flow is modelled as a frictional
dissipation process in continuum mechanics but the lack of influence of inter-particle friction on the energy
dissipation and spreading dynamics (Lube et al., 2005) is surprising.

\section*{Acknowledgements and References}
Acknowledgements and Reference heading should be left justified, bold, with the first letter capitalized but have no numbers. Text below continues as normal.

Online references will be linked to their original source, only if possible. To enable this linking extra care should be taken when preparing reference lists. References should be cited in the text by using \cite{RefJ}. A complete reference should provide enough information to locate the article.

%
% BibTeX or Biber users please use (the style is already called in the class, ensure 
% that the "woc.bst" style is in your local directory)
% \bibliography{name or your bibliography database}
%
% Non-BibTeX users please use
%
\begin{thebibliography}{}
%
% and use \bibitem to create references.
%
\bibitem{RefJ}
% Format for Journal Reference
Journal Author, Journal \textbf{Volume}, page numbers (year)
% Format for books
\bibitem{RefB}
Book Author, \textit{Book title} (Publisher, place, year) page numbers
% etc
\end{thebibliography}

\end{document}

% end of file template.tex

### 2016-11-20 paper.tex
Introduction and references

@@
 % Important: please activate and fill the "wocname" command with the exact title of the series for conferences not included in any of the series listed on the top
-%
-%\wocname{?????????}
-%
+\usepackage{siunitx}
+
 % Very important: please fill the "woctitle" command with the exact title of the conference
@@
   immersed granular flows, it is important to consider both the dynamics of the solid phase and the
-  role of the ambient fluid (Denlinger and Iverson, 2001). The dynamics of the solid phase alone are
+  role of the ambient fluid~\cite{Denlinger2001}. The dynamics of the solid phase alone are
   insufficient to describe the mechanism of granular flows in fluid. It is important to consider the
@@
   dense-compacted to dense-suspended flows, and the drag interactions which counteract the movement
-  of the solids (Meruane et al., 2010). Transient regimes characterised by a change in the solid
+  of the solids~\cite{Meruane2010}. Transient regimes characterised by a change in the solid
   fraction, dilation at the onset of flow and the development of excess pore-pressure, result in
@@
   vortices result in formation of heaps that significantly affect the distribution of mass in the flow.
-  Staron and Hinch (2007) observed that the distribution of mass in a granular flow plays a crucial role
+  ~\cite{Staron2007a} observed that the distribution of mass in a granular flow plays a crucial role
   in the flow kinematics. In order to understand the behaviour of tall columns, the run-out behaviour of a
@@
 the flow duration, the spreading velocity, the final extent of the deposit, and the energy dissipation can be scaled in a quantitative way independent of substrate properties, grain size, density, and shape of the granular
-material and released mass (Lajeunesse, Monnier, \& Homsy, 2005; Lube, Huppert, Sparks, \& Freundt, 2005).
-Simple mathematical models based on conservation of horizontal momentum capture the scaling laws of the
-final deposit, but fail to describe the initial transition regime. Granular flow is modelled as a frictional
-dissipation process in continuum mechanics but the lack of influence of inter-particle friction on the energy
-dissipation and spreading dynamics (Lube et al., 2005) is surprising.
+material and released mass~\cite{Lajeunesse2005, Lube2005}.
+Simple mathematical models based on conservation of horizontal momentum capture the scaling laws of the final deposit, but fail to describe the initial transition regime.
+Granular flow is modelled as a frictional dissipation process in continuum mechanics but the lack of influence of inter-particle friction on the energy dissipation and spreading dynamics~\cite{Lube2005} is surprising.
 
-\section*{Acknowledgements and References}
-Acknowledgements and Reference heading should be left justified, bold, with the first letter capitalized but have no numbers. Text below continues as normal.
-
-Online references will be linked to their original source, only if possible.
-To enable this linking extra care should be taken when preparing reference lists.
-References should be cited in the text by using \cite{RefJ}.
-A complete reference should provide enough information to locate the article.
+\section*{Acknowledgements}
+The author would like to thank the Cambridge Commonwealth and Overseas Trust for the financial support to the
+first author to pursue this research.
 
@@
 %
-\begin{thebibliography}{}
-%
-% and use \bibitem to create references.
-%
-\bibitem{RefJ}
-% Format for Journal Reference
-Journal Author, Journal \textbf{Volume}, page numbers (year)
-% Format for books
-\bibitem{RefB}
-Book Author, \textit{Book title} (Publisher, place, year) page numbers
-% etc
-\end{thebibliography}
+\bibliography{references}
 

### 2016-11-20 paper.tex
Tall columns

@@
 \usepackage{siunitx}
+\usepackage{cleveref}
+\usepackage{subcaption}
+\captionsetup{compatibility=false}
 
@@
   The effect of initial aspect ratio on the run-out behaviour is investigated. The flow kinematics are
-  compared with the dry and buoyant granular collapse to understand the influence of hydrodynamic forces
-  and lubrication on the run-out. In the case of tall columns, the amount of material destabilised above
-  the failure plane is larger than that of short columns. Hence in tall columns, the surface area of the
-  mobilised mass that interacts with the surrounding fluid is significantly higher than the short columns.
-  This increase in the area of soil - fluid interaction results in an increase in the formation of
-  turbulent vortices that alter the deposit morphology during the collapse. It is observed that the
-  vortices result in formation of heaps that significantly affect the distribution of mass in the flow.
-  ~\cite{Staron2007a} observed that the distribution of mass in a granular flow plays a crucial role
-  in the flow kinematics. In order to understand the behaviour of tall columns, the run-out behaviour of a
-  dense granular column with an initial aspect ratio of 6 is studied. The collapse of a tall granular
-  column on slopes of 0, 2.5, 5 and 7.5 are studied. The permeability of the granular mass was varied
-  to understand the effect of permeability on the run-out behaviour. }
+  compared with the dry and buoyant granular collapse to understand the influence of hydrodynamic
+  forces and lubrication on the run-out. In the case of tall columns, the amount of material
+  destabilised above the failure plane is larger than that of short columns. Hence in tall columns,
+  the surface area of the mobilised mass that interacts with the surrounding fluid is significantly
+  higher than the short columns. This increase in the area of soil - fluid interaction results in an
+  increase in the formation of turbulent vortices that alter the deposit morphology during the
+  collapse. It is observed that the vortices result in formation of heaps that significantly affect
+  the distribution of mass in the flow.~\cite{Staron2007a} observed that the distribution of mass in
+  a granular flow plays a crucial role in the flow kinematics. In order to understand the behaviour
+  of tall columns, the run-out behaviour of a dense granular column with an initial aspect ratio of
+  6 is studied. The collapse of a tall granular column on slopes of 0, 2.5, 5 and 7.5 are studied.
+  The permeability of the granular mass was varied to understand the effect of permeability on the
+  run-out behaviour. }
 %
@@
 \section{Introduction}
-The collapse of a granular column on a horizontal surface is a simple case of granular flow, however a
-proper model that describes the flow dynamics is still lacking. Experimental investigations have shown that
-the flow duration, the spreading velocity, the final extent of the deposit, and the energy dissipation can be scaled in a quantitative way independent of substrate properties, grain size, density, and shape of the granular
-material and released mass~\cite{Lajeunesse2005, Lube2005}.
-Simple mathematical models based on conservation of horizontal momentum capture the scaling laws of the final deposit, but fail to describe the initial transition regime.
-Granular flow is modelled as a frictional dissipation process in continuum mechanics but the lack of influence of inter-particle friction on the energy dissipation and spreading dynamics~\cite{Lube2005} is surprising.
+The collapse of a granular column on a horizontal surface is a simple case
+of granular flow, however a proper model that describes the flow dynamics
+is still lacking. Experimental investigations have shown that the flow
+duration, the spreading velocity, the final extent of the deposit, and the
+energy dissipation can be scaled in a quantitative way independent of
+substrate properties, grain size, density, and shape of the granular
+material and released mass~\cite{Lajeunesse2005, Lube2005}. Simple
+mathematical models based on conservation of horizontal momentum capture
+the scaling laws of the final deposit, but fail to describe the initial
+transition regime. Granular flow is modelled as a frictional dissipation
+process in continuum mechanics but the lack of influence of inter-particle
+friction on the energy dissipation and spreading dynamics~\cite{Lube2005}
+is surprising.
+
+\section{Tall columns}
+
+In the case of tall columns, the amount of material destabilised above the 
+failure plane is larger than that of short columns. Hence in tall columns, 
+the surface area of the mobilised mass that interacts with the 
+surrounding fluid is significantly higher than the short columns. This 
+increase in the area of soil - fluid interaction results in an increase
+in the formation of turbulent vortices that alter the deposit morphology during 
+the collapse. It is observed that the vortices result in formation of heaps 
+that significantly affect the distribution of mass in the flow.
+~\citet{Staron2007a} observed that the 
+distribution of mass in a granular flow plays a crucial role in the flow 
+kinematics. In order to understand the behaviour of tall columns, the run-out 
+behaviour of a dense granular column with an initial aspect ratio of 6 is 
+studied. The collapse of a tall granular column on slopes of 0\si{\degree}, 
+2.5\si{\degree}, 5\si{\degree} and 7.5\si{\degree} are studied. A hydrodynamic 
+radius of \textit{r} = 0.85 \textit{R} is adopted.
+
+Snapshots of the collapse of an aspect ratio 6 column in fluid on a horizontal 
+surface are shown in~\cref{fig:LBM_DEM_a6}. The initial stage of collapse is 
+characterised by the free-fall of grains above the failure surface. Unlike the 
+dry condition, as the grains undergo free-fall due to gravity, they interact 
+with the surrounding fluid experiencing drag forces. This results in a 
+significant drop in the kinetic energy available for the flow. As the grains 
+reach the static region, they interact with the neighbouring grains and the 
+kinetic energy gained during the free fall is converted into horizontal 
+acceleration. Uniquely during this stage ($t = 3\tau_c$) the interactions 
+between the soil grains on the surface with the surrounding fluid result in the 
+formation of eddies. The number of eddies formed during the flow is 
+proportional to the surface area of the granular mass interacting with the 
+fluid. Hydroplaning can be observed at the flow front ($t = 3\tau_c$). Two 
+large vortices with almost the same size can be observed at the final stage of 
+collapse. The soil grains on the surface experience suction due to formation of 
+eddies and this results in formation of heaps of granular mass in front of each 
+vortex. The formation of heaps, although in evidence, doesn't significantly 
+affect the distribution of mass in the case of collapse on a horizontal plane.
+
+\begin{figure}[htpb]
+\centering
+\includegraphics[width=0.9\linewidth]{figs/lbm_dem_a6}
+\caption{Flow evolution of a granular column collapse in fluid (a = 6) on a 
+horizontal surface.}
+\label{fig:LBM_DEM_a6}
+\end{figure}
+
+Snapshots of the collapse of a granular column (\textit{a} = 6) on a inclined 
+plane at angle of 5\si{\degree} are shown in~\cref{fig:a6_slope_snapshots}. 
+The collapse on a slope of 5\si{\degree} show flow evolution behaviour similar 
+to the case of collapse on a horizontal plane. The vortices are formed only 
+during the horizontal spreading stage $t = 3\tau_c$, but the number of vortices 
+formed during the collapse is higher than the collapse on a horizontal plane. 
+However, as the flow progresses a single large vortex engulfs other smaller 
+vortices, thus having a significant influence on the mass 
+distribution.~\Cref{fig:a6_slope_voro} shows 
+the distribution of mass and the packing density at $t = 6\tau_c$ and $t = 
+8\tau_c$. A heap can be observed in front of the large vortex almost at the 
+middle of the flow. The height of the heap formed in the middle of the granular 
+flow is higher than the collapse height next to the wall. However, when the 
+flow comes to rest and the vortex moves 
+away from the flowing surface, the mass present in the heap gets redistributed 
+(as seen at $t = 8\tau_c$). This behaviour is significantly different from that 
+observed in the case of short columns.
+
+\begin{figure}
+\makebox[\linewidth][c]{
+\begin{subfigure}[b]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_0}
+    \caption*{$t = 0\tau_c$}
+    \label{fig:a6slope5t_0}
+\end{subfigure}
+}\\
+
+\makebox[\linewidth][c]{
+\begin{subfigure}[b]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_1tc}
+    \caption*{$t = 1\tau_c$}
+    \label{fig:a6slope5t_1tc}
+\end{subfigure}
+}\\
+
+\makebox[\linewidth][c]{
+\begin{subfigure}[b]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_3tc}
+    \caption*{$t = 3\tau_c$}
+    \label{fig:a6slope5t_3tc}
+\end{subfigure}
+}\\
+
+\makebox[\linewidth][c]{
+\begin{subfigure}[b]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_6tc}
+    \caption*{$t = 6\tau_c$}
+    \label{fig:a6slope5t_6tc}
+\end{subfigure}
+}\\
+
+\makebox[\linewidth][c]{
+\begin{subfigure}[b]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_8tc}
+    \caption*{$t = 8\tau_c$}
+    \label{fig:a6slope5t_8tc}
+\end{subfigure}
+}
+\caption[Flow evolution of a granular column collapse in fluid (a = 6) on a 
+slope of 5\si{\degree}]{Flow evolution of a granular column collapse in fluid 
+(a = 6) on a slope of 5\si{\degree}. Shows the velocity profile of fluid due to 
+interaction with the grains (red - higher velocity).}
+\label{fig:a6_slope_snapshots}
+\end{figure}
+
+\begin{figure}
+\makebox[\linewidth][c]{
+\begin{subfigure}[b]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.95\linewidth]{figs/a6/a6_slope5_6tc}
+    \caption*{$t = 6\tau_c$}
+    \label{fig:a6_slope5_6tc}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[b]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.95\linewidth]{figs/a6/a6_slope5_8tc}
+    \caption*{$t = 8\tau_c$}
+    \label{fig:a6_slope5_8tc}
+\end{subfigure}
+}
+\caption{Packing density of a granular column collapse in fluid (a = 6) on a 
+slope of 5\si{\degree}.}
+\label{fig:a6_slope_voro}
+\end{figure}
+
+In order to understand the influence of slope angles on the run-out behaviour, 
+the collapse of a granular column with an initial aspect ratio of 6 is 
+performed on slopes of 0\si{\degree}, 2.5\si{\degree}, 5\si{\degree} and 
+7.5\si{\degree}. The run-out evolution with time for different slope angles are 
+presented in~\cref{fig:Runout_a6_slope}. The run-out distance increases with 
+increase in the slope angle, however the run-out distance in the fluid is 
+significantly shorter than the dry condition. The slow evolution of run-out in 
+the submerged condition is due to the delay in the dissipation of large 
+negative pore-pressures developed during the initial stage of the collapse. The 
+formation of eddies during the flow indicates that most of the potential energy 
+gained during the free-fall is dissipated through viscous drag and turbulence. 
+This effect predominates over the hydroplaning that is observed during 
+the flow resulting in a shorter run-out distance in the case of fluid. The 
+evolution of the normalised height with time (\cref{fig:Height_a6_slope}) for 
+collapse on different slope angles indicates that the 
+amount of material destabilised in fluid is less than the dry conditions due to 
+the drag forces experienced by the grains, which retards the quantity and the 
+rate of collapse.
+
+\begin{figure}
+\centering
+\begin{subfigure}[t]{0.9\linewidth}
+\includegraphics[width=0.9\linewidth]{figs/runout_a6_slope}
+\caption{Evolution of run-out for a column collapse in fluid (a = 6).}
+\label{fig:Runout_a6_slope}
+\end{subfigure} \\
+\begin{subfigure}[t]{0.9\linewidth}
+\centering
+\includegraphics[width=0.9\linewidth]{figs/height_a6_slope}
+\caption{Evolution of height with time for a column collapse in fluid (a = 6).}
+\label{fig:Height_a6_slope}
+\end{subfigure}
+\caption{Evolution of  run-out and height  for a column collapse in fluid (a = 
+6).}
+\label{fig:a6_slope}
+\end{figure}
+
+\Cref{fig:KE_a6_slope} shows the evolution of the normalised kinetic energy 
+for a granular column (a = 6) collapse in fluid on different slope angles. The 
+amount of kinetic energy available for the flow in the submerged condition 
+is almost half that of the dry condition. It can be seen from the figure that 
+the vertical kinetic energy in the fluid condition dissipates a longer 
+duration, in contrast to the free-fall release observed in the dry condition. 
+The slower dissipation is attributed to the viscous drag force experienced by 
+the grains.
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.8\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/KE_a6_slope}
+    \caption{Evolution of the total kinetic energy.}
+    \label{fig:KE_a6_slope}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/PE_a6_slope}
+    \caption{Evolution of the total potential energy.}
+    \label{fig:PE_a6_slope}
+\end{subfigure}
+}
+\caption{Evolution of the kinetic and the potential energy with time for a 
+granular column collapse in fluid (a = 6).}
+\label{fig:a6_KEPE}
+\end{figure}
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.8\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/KEx_a6_slope}
+    \caption{Evolution of the horizontal kinetic energy.}
+    \label{fig:KEx_a6_slope}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/KEy_a6_slope}
+    \caption{Evolution of the vertical kinetic energy.}
+    \label{fig:KEy_a6_slope}
+\end{subfigure}
+}
+\caption{Evolution of the kinetic energies with time for a granular 
+column collapse in fluid (a = 6).}
+\label{fig:a6_KExy}
+\end{figure}
+
+The behaviour of tall columns is significantly different from that observed in 
+the case of short columns. The slope angle has a strong influence on the number 
+and size of eddies during the flow. The eddies interact with the surface of the 
+granular flow and forms heaps in front of each vortex. This significantly 
+affects the mass distribution and in turn the run-out evolution. Although tall 
+cliffs are quite rare in submarine condition in comparison to short cliffs or 
+slopes, further research is required to understand the influence of 
+permeability and packing density on the run-out evolution of tall columns.
 

### 2016-11-21 paper.tex
Abstract and introduction

@@
         \firstname{Kenichi} \lastname{Soga}\inst{2}\fnsep\thanks{\email{soga@berkeley.edu}} \and
-        \firstname{Jean-Yves} \lastname{Delenne}\inst{3}\fnsep\thanks{\email{}} \and
-        \firstname{Farhang} \lastname{Radjai}\inst{4}\fnsep\thanks{\email{}}
-        % etc.
+        \firstname{Jean-Yves} \lastname{Delenne}\inst{3}\fnsep\thanks{\email{}}
 }
@@
 \and
-           University of California, Berkeley, US
+           University of California, Berkeley, USA
 \and
            INRA, France
-\and 
-           CNRS, France
 }
@@
 \abstract{%
-  Avalanches, landslides, and debris flows are geophysical hazards, which involve rapid mass movement
-  of granular solids, water, and air as a single phase system. In order to describe the mechanism of
-  immersed granular flows, it is important to consider both the dynamics of the solid phase and the
-  role of the ambient fluid~\cite{Denlinger2001}. The dynamics of the solid phase alone are
-  insufficient to describe the mechanism of granular flows in fluid. It is important to consider the
-  effect of hydrodynamic forces that reduce the weight of the solids inducing a transition from
-  dense-compacted to dense-suspended flows, and the drag interactions which counteract the movement
-  of the solids~\cite{Meruane2010}. Transient regimes characterised by a change in the solid
-  fraction, dilation at the onset of flow and the development of excess pore-pressure, result in
-  altering the balance between the stress carried by the fluid and that carried by the grains, thereby
-  changing the overall behaviour of the flow. In the present study, a two-dimensional coupled Lattice
-  Boltzmann LBM – DEM technique is developed to understand the micro-scale rheology of granular
-  flows in fluid.
-
-  In the present study, the collapse of a granular column in fluid is studied using 2D LBM - DEM.
-  The effect of initial aspect ratio on the run-out behaviour is investigated. The flow kinematics are
-  compared with the dry and buoyant granular collapse to understand the influence of hydrodynamic
-  forces and lubrication on the run-out. In the case of tall columns, the amount of material
-  destabilised above the failure plane is larger than that of short columns. Hence in tall columns,
-  the surface area of the mobilised mass that interacts with the surrounding fluid is significantly
-  higher than the short columns. This increase in the area of soil - fluid interaction results in an
-  increase in the formation of turbulent vortices that alter the deposit morphology during the
-  collapse. It is observed that the vortices result in formation of heaps that significantly affect
-  the distribution of mass in the flow.~\cite{Staron2007a} observed that the distribution of mass in
-  a granular flow plays a crucial role in the flow kinematics. In order to understand the behaviour
-  of tall columns, the run-out behaviour of a dense granular column with an initial aspect ratio of
-  6 is studied. The collapse of a tall granular column on slopes of 0, 2.5, 5 and 7.5 are studied.
-  The permeability of the granular mass was varied to understand the effect of permeability on the
-  run-out behaviour. }
+  Avalanches, landslides, and debris flows are geophysical hazards, involve
+  rapid mass movement of granular solids, water, and air as a multi-phase
+  system. In order to describe the mechanism of immersed granular flows,
+  it is important to consider both the dynamics of the solid phase and the
+  role of the ambient fluid~\cite{Denlinger2001}. In the present study,
+  the collapse of a granular column in fluid is
+  studied using 2D LBM - DEM. The flow kinematics are compared with the dry
+  and buoyant granular collapse to understand the influence of hydrodynamic
+  forces and lubrication on the run-out. In the case of tall columns,
+  the amount of material destabilised above the failure plane is larger than
+  that of short columns. Hence in tall columns, the surface area of the
+  mobilised mass that interacts with the surrounding fluid is significantly
+  higher than the short columns. This increase in the area of soil - fluid
+  interaction results in an increase in the formation of turbulent vortices
+  that alter the deposit morphology during the collapse. It is observed that
+  the vortices result in the formation of heaps that significantly affect
+  the distribution of mass in the flow. In order to understand the behaviour
+  of tall columns, the run-out behaviour
+  of a dense granular column with an initial aspect ratio of 6 is studied.
+  The collapse of a tall granular column on slopes of 0, 2.5, 5 and 7.5
+  are studied.}
 %
@@
 %
-\section{Introduction}
+\section{Collapse of granular columns}
 The collapse of a granular column on a horizontal surface is a simple case
@@
 substrate properties, grain size, density, and shape of the granular
-material and released mass~\cite{Lajeunesse2005, Lube2005}. Simple
-mathematical models based on conservation of horizontal momentum capture
-the scaling laws of the final deposit, but fail to describe the initial
-transition regime. Granular flow is modelled as a frictional dissipation
+material and released mass~\cite{Staron2007a, Lajeunesse2005, Lube2005}.
+
+Granular collapse on a horizontal plane exhibit two distinct flow regimes:
+(a) for columns with aspect ratio ‘a’ <1.7 a linear relation between the
+spread and aspect ratio can be observed, and (b) for ‘a’ > 1.7 a power-law
+relationship exists.~\citet{Soundararajan2015} observed that a simple
+frictional dissipation model is able to capture the flow dynamics for
+columns with small aspect ratios. However, for tall columns the flow
+dyanmics of is controlled by the initial collisional regime, resulting in a
+power-law relationship with the aspect ratios. Simple mathematical models
+based on conservation of horizontal momentum capture the scaling laws of
+the final deposit, but fail to describe the initial transition regime
+~\citep{Kumar2013}. Granular flow is modelled as a frictional dissipation
 process in continuum mechanics but the lack of influence of inter-particle
-friction on the energy dissipation and spreading dynamics~\cite{Lube2005}
-is surprising.
+collision and friction on the energy dissipation and spreading dynamics is
+surprising.
 

### 2016-11-21 paper.tex
Collapse on slope 5

@@
 substrate properties, grain size, density, and shape of the granular
-material and released mass~\cite{Staron2007a, Lajeunesse2005, Lube2005}.
+material and released mass
+~\cite{Topin2011, Staron2007a, Lajeunesse2005, Lube2005}.
 
 Granular collapse on a horizontal plane exhibit two distinct flow regimes:
-(a) for columns with aspect ratio ‘a’ <1.7 a linear relation between the
+(a) for columns with aspect ratio ‘a’ < 1.7 a linear relation between the
 spread and aspect ratio can be observed, and (b) for ‘a’ > 1.7 a power-law
-relationship exists.~\citet{Soundararajan2015} observed that a simple
-frictional dissipation model is able to capture the flow dynamics for
-columns with small aspect ratios. However, for tall columns the flow
+relationship exists. A simple frictional dissipation model is able to
+capture the flow dynamics for columns with small aspect ratios
+~\cite{soundararajan2015}. However, for tall columns the flow
 dyanmics of is controlled by the initial collisional regime, resulting in a
@@
 
-\section{Tall columns}
-
-In the case of tall columns, the amount of material destabilised above the 
-failure plane is larger than that of short columns. Hence in tall columns, 
-the surface area of the mobilised mass that interacts with the 
-surrounding fluid is significantly higher than the short columns. This 
-increase in the area of soil - fluid interaction results in an increase
+The amount of material destabilised above the failure plane, in tall columns,
+is larger than that of short columns. This increase in the mobilised mass
+results in a signficant increase in the surface area of granular mass that
+interacts with the surrounding fluid. The increase in the area of soil -
+fluid interaction results in an increase
 in the formation of turbulent vortices that alter the deposit morphology during 
@@
 that significantly affect the distribution of mass in the flow.
-~\citet{Staron2007a} observed that the 
-distribution of mass in a granular flow plays a crucial role in the flow 
-kinematics. In order to understand the behaviour of tall columns, the run-out 
-behaviour of a dense granular column with an initial aspect ratio of 6 is 
-studied. The collapse of a tall granular column on slopes of 0\si{\degree}, 
-2.5\si{\degree}, 5\si{\degree} and 7.5\si{\degree} are studied. A hydrodynamic 
-radius of \textit{r} = 0.85 \textit{R} is adopted.
+The distribution of mass in a granular flow plays a crucial role in the
+flow kinematics~\cite{Staron2007a}. In order to understand the
+behaviour of tall columns, the run-out behaviour of a dense granular column
+with an initial aspect ratio of 6 is studied using Lattice Boltzmann - Discrete
+Element Method coupling.
 
+\section{LBM-DEM coupling}
+The Lattice Boltzmann Method (LBM) is an alternative approach to
+the classical Navier-Stokes solvers for fluid flow and works on an
+equidistant grid of cells, called lattice cells, which interact only with
+their direct neighbours~\cite{He1997}. The fluid domain is divided into
+a rectangular grid or lattice, with the same spacing ‘h’ in both the x-
+and the y-directions. The present study focuses on 2-D problems; hence,
+the D2Q9 momentum discretization is adopted, where the fluid particles at
+each node are allowed to move to their eight immediate neighbours
+with eight different velocities $\mathit{e_i} (\mathit{i}=1\,,\dots\,,8)$.
+The primary variables in LB formulation are called the \textit{fluid 
+density distribution functions}, $\mathit{f_i}$, each representing the probable 
+amount of fluid particles moving with the velocity $\mathit{e_i}$ along the 
+direction $\mathit{i}$ at each node. In the present study, Lattice-Boltzmann
+- Multiple Relaxation Time (LBM-MRT) is used for numerical stability.
+The nine eigen values of $\mathbf{S}$ are all between 0 and 2 so as to
+maintain linear stability and the separation of scales. In  this  study,
+the Large Eddy Simulation (LES) is adopted to solve for turbulent flow
+problems. The separation of scales is achieved by filtering of the
+Navier-Stokes equations, from which the resolved scales are directly
+obtained and unresolved scales are modelled by a one-parameter
+Smagorinski sub-grid methodology~\cite{Smagorinsky1963}.
+
+Lattice Boltzmann approach can accommodate large grain sizes and the
+interaction between the fluid and the moving grains can be modelled
+through relatively simple fluid – grain interface treatments. Further,
+employing the Discrete Element Method (DEM) to account for the
+grain – grain interaction naturally leads to a combined LB – DEM
+procedure~\cite{Kumar2012}. The Eulerian nature of the LBM formulation,
+together with the common explicit time step scheme of both LBM and DEM
+makes this coupling strategy an efficient numerical procedure for the
+simulation of grain – fluid systems. A no-slip boundary condition is
+used to simulate grain – fluid and grain – grain interactions. The
+solid grains inside the fluid are represented by lattice
+nodes. The discrete nature of lattice, results in a stepwise
+representation of the surfaces, which are circular, hence
+sufficiently small lattice spacing is adopted. In order to model flow
+through interconnected pore space, a reduction in radius (The
+hydrodynamic radius \textit{r}) is assumed during the
+LBM computation~\citep{soundararajan2015}.
+
+\section{Numerical set-up}
+In this study, a 2D poly-disperse system ($d_{max} / d_{min} = 1.8$) of
+circular discs in fluid was used to understand the behaviour of granular
+flows on slopes (see~\cref{fig:geometry}.
+The soil column was modelled using 2500 discs of density
+2650~\si{\kilogram\per\meter\cubed} and a contact friction angle of
+26\si{\degree}.  The  collapse of the column was simulated inside a fluid
+with a density of 1000~\si{\kilogram\per\meter\cubed} and a kinematic viscosity
+of $1 \times 10^{-6}$~\si{\meter\squared\per\second}. The fluid was modelled
+with 5 to 9 million LB nodes. GPU computing was used to simulate the problem.
+The choice of a 2D geometry has the advantage of cheaper computational
+effort than a 3D case, makeing it feasible to simulate very large systems.
+A granular column of aspect ratio `\textit{a}' of 6 was used. A hydrodynamic
+radius of \textit{r} = 0.85 \textit{R} is adopted. Dry analyses were also
+performed to study the effect of hydrodynamic forces on the run-out distance.
+The collapse of a tall granular column on slopes of 0\si{\degree},
+2.5\si{\degree}, 5\si{\degree} and 7.5\si{\degree} are studied.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.95\linewidth]{figs/geometry}
+  \caption{Underwater granular collapse set-up}
+  \label{fig:geometry}
+\end{figure}
+
+\section{Collapse on a horizontal plane}
 Snapshots of the collapse of an aspect ratio 6 column in fluid on a horizontal 
@@
 
+\section{Collapse on slopes}
 Snapshots of the collapse of a granular column (\textit{a} = 6) on a inclined 
@@
 
-\begin{figure}
-\makebox[\linewidth][c]{
-\begin{subfigure}[b]{0.95\linewidth}
+\begin{figure}[htpb]
 \centering
-    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_0}
-    \caption*{$t = 0\tau_c$}
-    \label{fig:a6slope5t_0}
-\end{subfigure}
-}\\
-
-\makebox[\linewidth][c]{
-\begin{subfigure}[b]{0.95\linewidth}
-	\centering
-    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_1tc}
-    \caption*{$t = 1\tau_c$}
-    \label{fig:a6slope5t_1tc}
-\end{subfigure}
-}\\
-
-\makebox[\linewidth][c]{
-\begin{subfigure}[b]{0.95\linewidth}
-	\centering
-    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_3tc}
-    \caption*{$t = 3\tau_c$}
-    \label{fig:a6slope5t_3tc}
-\end{subfigure}
-}\\
-
-\makebox[\linewidth][c]{
-\begin{subfigure}[b]{0.95\linewidth}
-	\centering
-    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_6tc}
-    \caption*{$t = 6\tau_c$}
-    \label{fig:a6slope5t_6tc}
-\end{subfigure}
-}\\
-
-\makebox[\linewidth][c]{
-\begin{subfigure}[b]{0.95\linewidth}
-	\centering
-    \includegraphics[width=0.95\linewidth]{figs/a6/a6slope5t_8tc}
-    \caption*{$t = 8\tau_c$}
-    \label{fig:a6slope5t_8tc}
-\end{subfigure}
-}
+\includegraphics[width=0.9\linewidth]{figs/lbm_dem_a6_slope5}
 \caption[Flow evolution of a granular column collapse in fluid (a = 6) on a 
@@
 	\centering
-    \includegraphics[width=0.95\linewidth]{figs/a6/a6_slope5_6tc}
+    \includegraphics[width=0.95\linewidth]{figs/a6_slope5_6tc}
     \caption*{$t = 6\tau_c$}
@@
 	\centering
-    \includegraphics[width=0.95\linewidth]{figs/a6/a6_slope5_8tc}
+    \includegraphics[width=0.95\linewidth]{figs/a6_slope5_8tc}
     \caption*{$t = 8\tau_c$}
@@
 
+\section{Summary}
 The behaviour of tall columns is significantly different from that observed in 

### 2016-11-22 paper.tex
data
Completed draft

@@
   it is important to consider both the dynamics of the solid phase and the
-  role of the ambient fluid~\cite{Denlinger2001}. In the present study,
+  role of the ambient fluid. In the present study,
   the collapse of a granular column in fluid is
@@
 the final deposit, but fail to describe the initial transition regime
-~\citep{Kumar2013}. Granular flow is modelled as a frictional dissipation
+~\citep{Kumar2012}. Granular flow is modelled as a frictional dissipation
 process in continuum mechanics but the lack of influence of inter-particle
@@
 obtained and unresolved scales are modelled by a one-parameter
-Smagorinski sub-grid methodology~\cite{Smagorinsky1963}.
+Smagorinski sub-grid methodology.%~\cite{Smagorinsky1963}.
 
@@
 surface are shown in~\cref{fig:LBM_DEM_a6}. The initial stage of collapse is 
-characterised by the free-fall of grains above the failure surface. Unlike the 
-dry condition, as the grains undergo free-fall due to gravity, they interact 
+characterised by the free-fall of grains above the failure surface.
+The height of the static region, which is below the shear-failure surface,
+is shorter than the total height of the column. As the grains experience
+free-fall they interact with the surrounding fluid. Unlike the dry condition,
+as the grains undergo free-fall due to gravity, they interact 
 with the surrounding fluid experiencing drag forces. This results in a 
-significant drop in the kinetic energy available for the flow. As the grains 
-reach the static region, they interact with the neighbouring grains and the 
-kinetic energy gained during the free fall is converted into horizontal 
-acceleration. Uniquely during this stage ($t = 3\tau_c$) the interactions 
-between the soil grains on the surface with the surrounding fluid result in the 
-formation of eddies. The number of eddies formed during the flow is 
-proportional to the surface area of the granular mass interacting with the 
-fluid. Hydroplaning can be observed at the flow front ($t = 3\tau_c$). Two 
-large vortices with almost the same size can be observed at the final stage of 
-collapse. The soil grains on the surface experience suction due to formation of 
-eddies and this results in formation of heaps of granular mass in front of each 
-vortex. The formation of heaps, although in evidence, doesn't significantly 
-affect the distribution of mass in the case of collapse on a horizontal plane.
+significant drop in the kinetic energy available for the flow.
+However, no vortices are observed during the initial stage of collapse.
 
@@
 
+As the grains reach the static region, they interact with the neighbouring
+grains and the kinetic energy gained during the free fall is converted into
+horizontal acceleration. Uniquely during this stage ($t = 3\tau_c$) the
+interactions between the soil grains on the surface with the surrounding
+fluid result in the formation of eddies. The number of eddies formed during
+the flow is proportional to the surface area of the granular mass interacting
+with the fluid. Hydroplaning can be observed at the flow front ($t = 3\tau_c$).
+Two large vortices with almost the same size can be observed at the final stage of 
+collapse. The soil grains on the surface experience suction due to formation of 
+eddies and this results in formation of heaps of granular mass in front of each 
+vortex. Unlike short columns, these vortices have 
+a significant influence on the mass distribution along the run-out. Heaps of 
+granular material can be observed in front of each vortex. The number of 
+vortices formed during a collapse is found to be proportional to the amount of 
+material destabilised, i.e., the length of free-surface interacting with the 
+fluid influences the number of vortices generated during the collapse. The 
+reappearance of force chains at $t = 6\tau_c$ indicates the 
+granular mass is consolidating resulting in an increase in the shear strength. 
+The formation of heaps, although in evidence, doesn't significantly 
+affect the distribution of mass in the case of collapse on a horizontal plane.
+
 \section{Collapse on slopes}
@@
 during the horizontal spreading stage $t = 3\tau_c$, but the number of vortices 
-formed during the collapse is higher than the collapse on a horizontal plane. 
+formed during the collapse is more than the collapse on a horizontal plane. 
 However, as the flow progresses a single large vortex engulfs other smaller 
 vortices, thus having a significant influence on the mass 
-distribution.~\Cref{fig:a6_slope_voro} shows 
-the distribution of mass and the packing density at $t = 6\tau_c$ and $t = 
-8\tau_c$. A heap can be observed in front of the large vortex almost at the 
-middle of the flow. The height of the heap formed in the middle of the granular 
-flow is higher than the collapse height next to the wall. However, when the 
-flow comes to rest and the vortex moves 
-away from the flowing surface, the mass present in the heap gets redistributed 
-(as seen at $t = 8\tau_c$). This behaviour is significantly different from that 
-observed in the case of short columns.
+distribution.
 
-\begin{figure}[htpb]
+\begin{figure}[!h]
 \centering
@@
 \end{figure}
+
+~\Cref{fig:a6_slope_voro} shows 
+the distribution of mass and the packing density at $t = 6\tau_c$ and
+$t = 8\tau_c$. A heap can be observed in front of the large vortex almost at the 
+middle of the flow. The height of the heap formed in the middle of the granular 
+flow is higher than the collapse height next to the wall. However, when the 
+flow comes to rest and the vortex moves 
+away from the flowing surface, the mass present in the heap gets redistributed 
+(as seen at $t = 8\tau_c$). This behaviour is significantly different from that 
+observed in the case of short columns. More denser packing can be observed in front
+of the eddies, resulting in spectrums dense - loose regions.
 
@@
 
+\subsection{Effect of slope angle}
 In order to understand the influence of slope angles on the run-out behaviour, 
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 \centering
-\begin{subfigure}[t]{0.9\linewidth}
 \includegraphics[width=0.9\linewidth]{figs/runout_a6_slope}
@@
 \label{fig:Runout_a6_slope}
-\end{subfigure} \\
-\begin{subfigure}[t]{0.9\linewidth}
+\end{figure}
+
+\begin{figure}[htbp]
 \centering
-\includegraphics[width=0.9\linewidth]{figs/height_a6_slope}
+\includegraphics[width=0.8\linewidth]{figs/height_a6_slope}
 \caption{Evolution of height with time for a column collapse in fluid (a = 6).}
 \label{fig:Height_a6_slope}
-\end{subfigure}
-\caption{Evolution of  run-out and height  for a column collapse in fluid (a = 
-6).}
-\label{fig:a6_slope}
 \end{figure}
 
-\Cref{fig:KE_a6_slope} shows the evolution of the normalised kinetic energy 
-for a granular column (a = 6) collapse in fluid on different slope angles. The 
+\Cref{fig:KEx_a6_slope,fig:KEy_a6_slope} shows the evolution of the normalised
+kinetic energies on different slope angles. The 
 amount of kinetic energy available for the flow in the submerged condition 
-is almost half that of the dry condition. It can be seen from the figure that 
-the vertical kinetic energy in the fluid condition dissipates a longer 
-duration, in contrast to the free-fall release observed in the dry condition. 
-The slower dissipation is attributed to the viscous drag force experienced by 
-the grains.
+is almost half that of the dry condition.
 
-\begin{figure}
+\Cref{fig:KEx_a6_slope} shows that in the dry condition, the kinetic energy
+available for the flow increases proportionally with the slope angle. However,
+in the case of submerged flow slopes 0\si{\degree} and 2\si{\degree},
+and 5\si{\degree} and 7.5\si{\degree} have similar kinetic energy evolution.
+The longer tail observed in the submerged granular collapse can be attributed to
+the formation and migration of eddies and subsequent mass redistribution into heaps.
+
+\begin{figure}[htbp]
 \centering
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.8\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/KE_a6_slope}
-    \caption{Evolution of the total kinetic energy.}
-    \label{fig:KE_a6_slope}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.95\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/PE_a6_slope}
-    \caption{Evolution of the total potential energy.}
-    \label{fig:PE_a6_slope}
-\end{subfigure}
-}
-\caption{Evolution of the kinetic and the potential energy with time for a 
-granular column collapse in fluid (a = 6).}
-\label{fig:a6_KEPE}
+\includegraphics[width=0.9\linewidth]{figs/KEx_a6_slope}
+\caption{Evolution of horizontal kinetic energy with time for a granular 
+column collapse in fluid (a = 6).}
+\label{fig:KEx_a6_slope}
 \end{figure}
 
-\begin{figure}
+\Cref{fig:KEx_a6_slope} shows that 
+the vertical kinetic energy in the fluid condition dissipates slower, in
+comparison to the free-fall release observed in the dry condition. The slower
+dissipation is attributed to the viscous drag force experienced by the grains. 
+However, the slope angle has negligble effect on the evolution of vertical
+kinetic energy in submerged condition. Where as, dry conditions show 
+$\approx15\%$ increase in the peak vertical kinetic energy when the slope angle
+changes from 0\si{\degree} to 7.5\si{\degree}.
+
+\begin{figure}[htbp]
 \centering
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.8\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/KEx_a6_slope}
-    \caption{Evolution of the horizontal kinetic energy.}
-    \label{fig:KEx_a6_slope}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.95\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/KEy_a6_slope}
-    \caption{Evolution of the vertical kinetic energy.}
+\includegraphics[width=0.9\linewidth]{figs/KEy_a6_slope}
+\caption{Evolution of vertical kinetic energy with time for a granular 
+column collapse in fluid (a = 6).}
 \label{fig:KEy_a6_slope}
-\end{subfigure}
-}
-\caption{Evolution of the kinetic energies with time for a granular 
-column collapse in fluid (a = 6).}
-\label{fig:a6_KExy}
 \end{figure}
@@
 The behaviour of tall columns is significantly different from that observed in 
-the case of short columns. The slope angle has a strong influence on the number 
+short columns. The slope angle has a strong influence on the number 
 and size of eddies during the flow. The eddies interact with the surface of the 
@@
 affects the mass distribution and in turn the run-out evolution. Although tall 
-cliffs are quite rare in submarine condition in comparison to short cliffs or 
-slopes, further research is required to understand the influence of 
-permeability and packing density on the run-out evolution of tall columns.
+cliffs are quite rare in submarine condition, further research is
+required to understand the influence of permeability and packing density
+on the run-out behaviour.
 
 \section*{Acknowledgements}
-The author would like to thank the Cambridge Commonwealth and Overseas Trust for the financial support to the
-first author to pursue this research.
+The author would like to thank the Cambridge Commonwealth and Overseas Trust
+for the financial support to pursue this research.
 

### 2016-11-22 paper.tex
Version 1.0

@@
         \firstname{Kenichi} \lastname{Soga}\inst{2}\fnsep\thanks{\email{soga@berkeley.edu}} \and
-        \firstname{Jean-Yves} \lastname{Delenne}\inst{3}\fnsep\thanks{\email{}}
+        \firstname{Jean-Yves} \lastname{Delenne}\inst{3}\fnsep\thanks{\email{delenne@supagro.inra.fr}}
 }
@@
 \and
-           INRA, France
+            INRA, CIRAD, University Montpellier 2 and SupAgro, France
 }
@@
 ~\cite{soundararajan2015}. However, for tall columns the flow
-dyanmics of is controlled by the initial collisional regime, resulting in a
+dyanmics is controlled by the initial collisional regime, resulting in a
 power-law relationship with the aspect ratios. Simple mathematical models
@@
 the final deposit, but fail to describe the initial transition regime
-~\citep{Kumar2012}. Granular flow is modelled as a frictional dissipation
+~\cite{Kumar2012}. Granular flow is modelled as a frictional dissipation
 process in continuum mechanics but the lack of influence of inter-particle
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figs/KEy_a6_slope}
+\includegraphics[width=0.8\linewidth]{figs/KEy_a6_slope}
 \caption{Evolution of vertical kinetic energy with time for a granular 

### 2017-02-07 paper.tex
Update figures to say fluid velocity

@@
 circular discs in fluid was used to understand the behaviour of granular
-flows on slopes (see~\cref{fig:geometry}.
+flows on slopes (see~\cref{fig:geometry}).
 The soil column was modelled using 2500 discs of density
@@
   \includegraphics[width=0.95\linewidth]{figs/geometry}
-  \caption{Underwater granular collapse set-up}
+  \caption{Underwater granular collapse set-up.
+    The slope angle is represented by $\theta$.}
   \label{fig:geometry}

### 2017-02-07 paper.tex
Changed long sentence

@@
 2650~\si{\kilogram\per\meter\cubed} and a contact friction angle of
-26\si{\degree}.  The  collapse of the column was simulated inside a fluid
+26\si{\degree} and a mean packing fraction of $\approx 0.82$.
+The  collapse of the column was simulated inside a fluid
 with a density of 1000~\si{\kilogram\per\meter\cubed} and a kinematic viscosity
@@
 collapse on different slope angles indicates that the 
-amount of material destabilised in fluid is less than the dry conditions due to 
-the drag forces experienced by the grains, which retards the quantity and the 
-rate of collapse.
+amount of material destabilised in fluid is less than the dry conditions.
+This is due to the drag forces experienced by the grains, which retards the
+quantity and the rate of collapse.
 

### 2017-02-13 paper.tex
Minor corrections

@@
 \usepackage{subcaption}
+\usepackage{gensymb}
 \captionsetup{compatibility=false}
@@
 \begin{document}
+\setlength{\belowcaptionskip}{-5pt}
+
 %
@@
 \abstract{%
-  Avalanches, landslides, and debris flows are geophysical hazards, involve
+  Avalanches, landslides, and debris flows are geophysical hazards, which involve
   rapid mass movement of granular solids, water, and air as a multi-phase
@@
   the amount of material destabilised above the failure plane is larger than
-  that of short columns. Hence in tall columns, the surface area of the
-  mobilised mass that interacts with the surrounding fluid is significantly
+  that of short columns. Therefore, the surface area of the
+  mobilised mass that interacts with the surrounding fluid in tall columns is significantly
   higher than the short columns. This increase in the area of soil - fluid
   interaction results in an increase in the formation of turbulent vortices
-  that alter the deposit morphology during the collapse. It is observed that
-  the vortices result in the formation of heaps that significantly affect
+  thereby altering the deposit morphology. It is observed that
+  the vortices result in the formation of heaps that significantly affects
   the distribution of mass in the flow. In order to understand the behaviour
@@
   of a dense granular column with an initial aspect ratio of 6 is studied.
-  The collapse of a tall granular column on slopes of 0, 2.5, 5 and 7.5
-  are studied.}
+  The collapse behaviour is analysed for different slope angles: 0\si{\degree}, 2.5\si{\degree}, 5\si{\degree} and 7.5\si{\degree}.}
 %
@@
 The collapse of a granular column on a horizontal surface is a simple case
-of granular flow, however a proper model that describes the flow dynamics
+of granular flow, however, a proper model that describes the flow dynamics
 is still lacking. Experimental investigations have shown that the flow
@@
 energy dissipation can be scaled in a quantitative way independent of
-substrate properties, grain size, density, and shape of the granular
+substrate properties, grain size, packing density, and shape of the granular
 material and released mass
@@
 
-Granular collapse on a horizontal plane exhibit two distinct flow regimes:
-(a) for columns with aspect ratio ‘a’ < 1.7 a linear relation between the
-spread and aspect ratio can be observed, and (b) for ‘a’ > 1.7 a power-law
+Collapse of granular columns on a horizontal plane exhibit two distinct flow regimes:
+(a) for columns with an initial aspect ratio `\textit{a}’ < 1.7, a linear relation between the
+spread and aspect ratio can be observed, and (b) for `\textit{a}’ > 1.7, a power-law
 relationship exists. A simple frictional dissipation model is able to
@@
 ~\cite{soundararajan2015}. However, for tall columns the flow
-dyanmics is controlled by the initial collisional regime, resulting in a
-power-law relationship with the aspect ratios. Simple mathematical models
-based on conservation of horizontal momentum capture the scaling laws of
-the final deposit, but fail to describe the initial transition regime
-~\cite{Kumar2012}. Granular flow is modelled as a frictional dissipation
-process in continuum mechanics but the lack of influence of inter-particle
-collision and friction on the energy dissipation and spreading dynamics is
-surprising.
+dynamics is controlled by the initial collisional regime, resulting in a
+power-law relationship with the initial aspect ratio of the columns. 
+Simple mathematical models based on conservation of horizontal momentum 
+capture the scaling laws of the final deposit, but fail to describe the 
+initial transition regime ~\cite{Kumar2012}. Granular flow is modelled as 
+a frictional dissipation process in continuum mechanics but the lack of 
+influence of inter-particle collision and friction on the energy dissipation and spreading dynamics is surprising.
 
@@
 is larger than that of short columns. This increase in the mobilised mass
-results in a signficant increase in the surface area of granular mass that
+results in a significant increase in the surface area of granular mass that
 interacts with the surrounding fluid. The increase in the area of soil -
 fluid interaction results in an increase
-in the formation of turbulent vortices that alter the deposit morphology during 
-the collapse. It is observed that the vortices result in formation of heaps 
-that significantly affect the distribution of mass in the flow.
+in the formation of turbulent vortices, which alter the deposit morphology. 
+It is observed that the vortices result in formation of heaps 
+thereby redistributing the granular mass in the flow.
 The distribution of mass in a granular flow plays a crucial role in the
@@
 their direct neighbours~\cite{He1997}. The fluid domain is divided into
-a rectangular grid or lattice, with the same spacing ‘h’ in both the x-
-and the y-directions. The present study focuses on 2-D problems; hence,
-the D2Q9 momentum discretization is adopted, where the fluid particles at
+a rectangular grid or lattice, with the same spacing `\textit{h}’ in both the \textit{x}-
+and the \textit{y}-directions. The present study focuses on 2-D problems; hence,
+the D2Q9 momentum discretization is adopted. The fluid particles at
 each node are allowed to move to their eight immediate neighbours
@@
 obtained and unresolved scales are modelled by a one-parameter
-Smagorinski sub-grid methodology.%~\cite{Smagorinsky1963}.
+Smagorinski sub-grid methodology~\cite{Smagorinsky1963}.
 
@@
 with a density of 1000~\si{\kilogram\per\meter\cubed} and a kinematic viscosity
-of $1 \times 10^{-6}$~\si{\meter\squared\per\second}. The fluid was modelled
+of $1 \times 10^{-6}$~\si{\meter\squared\per\second}. The fluid domain was modelled
 with 5 to 9 million LB nodes. GPU computing was used to simulate the problem.
 The choice of a 2D geometry has the advantage of cheaper computational
-effort than a 3D case, makeing it feasible to simulate very large systems.
+effort than a 3D case, making it feasible to simulate very large systems.
 A granular column of aspect ratio `\textit{a}' of 6 was used. A hydrodynamic
@@
 observed in the case of short columns. More denser packing can be observed in front
-of the eddies, resulting in spectrums dense - loose regions.
+of the eddies, resulting in spectrum of dense - loose regions.
 
@@
 }
-\caption{Packing density of a granular column collapse in fluid (a = 6) on a 
-slope of 5\si{\degree}.}
+\caption{Packing density (black - maximum density) of a granular column collapse in fluid (a = 6) on a slope of 5\si{\degree}.}
 \label{fig:a6_slope_voro}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[tbph]
 \centering
-\includegraphics[width=0.9\linewidth]{figs/runout_a6_slope}
+\includegraphics[width=0.95\linewidth]{figs/runout_a6_slope}
 \caption{Evolution of run-out for a column collapse in fluid (a = 6).}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[tbph]
 \centering
-\includegraphics[width=0.8\linewidth]{figs/height_a6_slope}
+\includegraphics[width=\linewidth]{figs/height_a6_slope}
 \caption{Evolution of height with time for a column collapse in fluid (a = 6).}
@@
 
-\Cref{fig:KEx_a6_slope,fig:KEy_a6_slope} shows the evolution of the normalised
-kinetic energies on different slope angles. The 
+\Cref{fig:KEx_a6_slope} shows that in the dry condition, the kinetic energy
+available for the flow increases proportionally with the slope angle. The 
 amount of kinetic energy available for the flow in the submerged condition 
-is almost half that of the dry condition.
-
-\Cref{fig:KEx_a6_slope} shows that in the dry condition, the kinetic energy
-available for the flow increases proportionally with the slope angle. However,
-in the case of submerged flow slopes 0\si{\degree} and 2\si{\degree},
+is almost half that of the dry condition. However,
+in the case of submerged flow, slopes 0\si{\degree} and 2\si{\degree},
 and 5\si{\degree} and 7.5\si{\degree} have similar kinetic energy evolution.
@@
 
-\begin{figure}[htbp]
+\begin{figure}[tbph]
 \centering
-\includegraphics[width=0.9\linewidth]{figs/KEx_a6_slope}
+\includegraphics[width=\linewidth]{figs/KEx_a6_slope}
 \caption{Evolution of horizontal kinetic energy with time for a granular 
@@
 
-\Cref{fig:KEx_a6_slope} shows that 
+\Cref{fig:KEy_a6_slope} shows that 
 the vertical kinetic energy in the fluid condition dissipates slower, in
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!tbph]
 \centering
-\includegraphics[width=0.8\linewidth]{figs/KEy_a6_slope}
+\includegraphics[width=0.95\linewidth]{figs/KEy_a6_slope}
 \caption{Evolution of vertical kinetic energy with time for a granular 
@@
 
-%
-% BibTeX or Biber users please use (the style is already called in the class, ensure 
-% that the "woc.bst" style is in your local directory)
-% \bibliography{name or your bibliography database}
-%
-% Non-BibTeX users please use
-%
 \bibliography{references}
@@
 \end{document}
-
-% end of file template.tex

### 2017-02-13 reviewer_comments_reply.md
Reply to reviewer

# Referee 1:

```
The paper presents new results on the dynamics of tall granular columns that collapse in a fluid.
The text is clear yet some sentences are too long and difficult to understand (e.g. section 5.1:

The evolution of the normalised height with time (fig. 6) for collapse on different slope angles indicates that the amount of material destabilised in fluid is less than the dry conditions due to
the drag forces experienced by the grains, which retards the quantity and the rate of collapse.

For clearness you may break these long sentences into shorter sentences.
Figures are insightful yet some details may be improved:
```

The authors would like to than the reviewer for useful comments. The following changes have been made incorporating the changes suggested by the reviewers.

> Figure 1: define angle theta as the slope angle (?)

Slope angle theta is now defined in the figure.

> Figures 2, 3: Color scales seems to correspond to two independent parameters: velocity and grain pressure. Does this means that velocity scale only applies for fluid ? If so then write ‘fluid velocity’ instead of velocity; if not then please explain.

The suggested update to the contour legends has been made.

>Figure 5: explain the meaning of the gray scale color for particles (in terms of mass packing?).

Black colour denotes maximum density of packing, while white denotes loose packing. Explanation has been added to the figure.

```
The paper is suitable for publication once these modifications are completed.

The topic is suitable for an oral presentation showing animations of the simulations, yet a poster or poster presentation but for an oral presentation it is also acceptable.
```

# Referee 2:
```
The content of the coupling of DEM-LBM is interesting and the credible results are shown.
The following conmments and the information are needed as follows:
```
> How and why did the authors determine the parameters for the simulations?

The parameters of the simulations are based on a typical soil media. Further details are provided in the PhD thesis Soundararajan (2015).

>How did the authors controll the pore water pressure caused by the change in void sizes?

The pore-presssure in the void size is modelled using LB nodes. As the grains move, and the void ratio is changes, solid nodes transition to fluid nodes and the pore-pressure is modelled using LB technique.

>How much is the density of granular mass of the initial packing?

The density of packing fraction is 0.82 and has been added to the text.
